---
title: "Plotting Fixed Effects Solutions From EBV-Routine For MAR in BVCH"
date: "`r Sys.Date()`"
---

# Pre-Requisites
Required packages are loaded 

```{r setup-req-pkg, message=FALSE, warning=FALSE}
require(readr)
require(dplyr)
require(ggplot2)
```

# Disclaimer
Experiments and trials to plot fixed effect solutions from routine EBV for MAR 
in BVCH are shown in this document.

# Data
The data is located in the work directory of the routine EBV for MAR

```{r data-dir}
s_breed <- "bv"
s_trait <- "mar"
s_data_dir_pre <- "/qualstorzws01/data_zws/health"
s_data_dir_post <- "zws"
s_data_dir <- file.path(s_data_dir_pre, s_trait, "work", s_breed, s_data_dir_post)
# check whether data directory exists
if (!dir.exists(s_data_dir)) stop(" *** * ERROR: Cannot find data directory: ", s_data_dir)
```

## Effect Levels
Try to extract effect levels from the re-coding log file. Determine the 
path to the logfile. 

```{r extract-level}
s_log_fn <- "codeEff.log"
s_log_file_path <- file.path(s_data_dir, s_log_fn)
```

### Extractor Helper Function
For a given row in the logfile, the values for calvingYear (cy), calvingMonth (cm)
and the corresponding level are extracted and returned as a one-row tibble

```{r helper-extractor-function}
# extractor function to get from single record, the values for cy, cm and the level
tbl_extract_code_info <- function(pvec_log_info){
  # get index in pvec_log_info where cy info is
  vec_cy_idx <- grep(pattern = "calvingYear", pvec_log_info)
  # get cy info and convert to number
  n_cy <- as.integer(gsub(pattern = "cycm(calvingYear=", 
                          replacement = "", 
                          pvec_log_info[vec_cy_idx], 
                          fixed = TRUE))
  # get index for cm info
  vec_cm_idx <- grep(pattern = "calvingMonth", pvec_log_info)
  # get level code
  vec_cm_info <- unlist(strsplit(pvec_log_info[vec_cm_idx], split = ") ", fixed = TRUE))
  n_code <- as.integer(vec_cm_info[2])
  n_cm <- as.integer(gsub(pattern = "calvingMonth=", 
                          replacement = "", 
                          vec_cm_info[1],
                          fixed = TRUE))
  return(tibble::tibble(calvingYear = n_cy, calvingMonth = n_cm, level = n_code))
}
```

### Main Function producing the code-level mapping
The function below reads the logfile and extracts the part with the mapping. 
The different code levels are extracted from each line using the above helper 
function.

```{r extract-tbl-with-level-codes}
tbl_get_level_code <- function(ps_log_file_path) {
  # check whether log file exists
  if (!file.exists(ps_log_file_path)) stop(" *** Error: Cannot find log file: ", ps_log_file_path)
  # read logfile into vector of string
  con_code_eff_log <- file(ps_log_file_path)
  vec_code_eff_log <- readLines(con = con_code_eff_log)
  close(con = con_code_eff_log)
  # get idx where mapping of level-to-codes is
  vec_code_eff_idx <- grep(pattern = "cycm", vec_code_eff_log, fixed = TRUE)
  vec_code_eff_level <- vec_code_eff_log[vec_code_eff_idx]
  # split every line into entries for calvingYear and calvingMonth
  l_code_info <- strsplit(vec_code_eff_level, split = ", ", fixed = TRUE)

  # return the result
  return(purrr::map_dfr(l_code_info, tbl_extract_code_info))
  
}
```

### Code Extraction
The function `tbl_get_level_code` is used to extract the codes of the different fixed effect levels

```{r extract-code}
tbl_level_code <- tbl_get_level_code(ps_log_file_path = s_log_file_path)
# head(tbl_level_code)
```

### Limit Number of Years
The analysis is done for the last few years and therefore, the codes are only kept for these years

```{r restrict-code}
n_yr_back <- 5
n_min_year <- max(tbl_level_code$calvingYear) - n_yr_back
tbl_level_code <- tbl_level_code %>%
  filter(calvingYear > n_min_year)
# head(tbl_level_code)
# dim(tbl_level_code)
```

### Same Code for Multiple Levels
Some codes are used for multiple levels. This is verified in the following statement.

```{r multi-level-check}
tbl_mult_level_code <- tbl_level_code %>% 
  group_by(level) %>% 
  summarise(cnt = n()) %>% 
  filter(cnt > 1)
tbl_mult_level_code
```

At the moment we are not eliminating any level-code mappings.

```{r elim-level-code}
# eliminate duplicate code explicitly, this a manual step and can only be done once
# if (nrow(tbl_mult_level_code) > 0){
#   tbl_level_code %>% filter(level %in% tbl_mult_level_code$level)
#   tbl_level_code <- tbl_level_code[-1,]
#   
# }
```

## Reading The Solutions
The solution file from MiX99 for the trait MAR for BVCH is read

```{r read-sol-fix}
s_solfix_fn <- "Solfix"
s_solfix_path <- file.path(s_data_dir, s_solfix_fn)
if (!file.exists(s_solfix_path)) stop(" *** * ERROR: Cannot find solfix file: ", s_solfix_path)
tbl_sol_fix <- read_delim(file      = s_solfix_path, 
                          delim     = " ") 
```

Remove leading and trailing spaces from column names                 

```{r rem-spaces-from-colnames}
colnames(tbl_sol_fix) <- purrr::map(colnames(tbl_sol_fix), function(x) gsub(pattern = "\\s+", replacement = "", x))
#head(tbl_sol_fix)
```

Remove spaces from entries in the tibble.

```{r rem-spaces-from-tibble}
tbl_sol_fix <- purrr::modify(tbl_sol_fix,  function(x) gsub(pattern = "\\s+", replacement = "", x))
#head(tbl_sol_fix)
```

Data-types for the columns `Fact.`, `Trt`, `Level`, `N-Obs` and `Solution` are changed

```{r change-data-types}
tbl_sol_fix$Fact. <- as.integer(tbl_sol_fix$Fact.)
tbl_sol_fix$Trt <- as.integer(tbl_sol_fix$Trt)
tbl_sol_fix$Level <- as.integer(tbl_sol_fix$Level)
tbl_sol_fix$`N-Obs` <- as.integer(tbl_sol_fix$`N-Obs`)
tbl_sol_fix$Solution <- as.double(tbl_sol_fix$Solution)
#head(tbl_sol_fix)
```

The solutions read from the solution file are filtered such that only solutions 
for the trait `mar` and the effect `cycm` are kept.

```{r filter-trait-mar-factor-cycm}
tbl_sol_fix_mar_cycm <- tbl_sol_fix %>% 
  filter(Trait == 'mar' & Factor == 'cycm')
tbl_sol_fix_hsa150_cycm <- tbl_sol_fix %>% 
  filter(Trait == 'hsa150' & Factor == 'cycm')
tbl_sol_fix_ms150_cycm <- tbl_sol_fix %>% 
  filter(Trait == 'ms150' & Factor == 'cycm')

#head(tbl_sol_fix_mar_cycm)
```

## Merge The Levels To the Solutions Based on Code-Level-Mapping
The levels are required for plot as a meaningful x-Axis display

```{r merge-levels-to-codes}
tbl_sol_fix_mar_cycm <- tbl_sol_fix_mar_cycm %>%
  inner_join(tbl_level_code, by = c('Level' = 'level'))
#head(tbl_sol_fix_mar_cycm)
tbl_sol_fix_hsa150_cycm <- tbl_sol_fix_hsa150_cycm %>%
  inner_join(tbl_level_code, by = c('Level' = 'level'))
tbl_sol_fix_ms150_cycm <- tbl_sol_fix_ms150_cycm %>%
  inner_join(tbl_level_code, by = c('Level' = 'level'))
```

## Restrict Range of Calving Years
The analysis is only done for a given range of calving years. The data is filtered 
to keep only records within the desired range.

```{r restrict-calving-year}
n_min_year <- max(tbl_sol_fix_mar_cycm$calvingYear) - n_yr_back
n_max_year <- 2018
tbl_sol_fix_mar_cycm <- tbl_sol_fix_mar_cycm %>%
  filter(calvingYear > n_min_year) %>%
  filter(calvingYear <= n_max_year)
#head(tbl_sol_fix_mar_cycm)
tbl_sol_fix_hsa150_cycm <- tbl_sol_fix_hsa150_cycm %>%
  filter(calvingYear > n_min_year) %>%
  filter(calvingYear <= n_max_year)
tbl_sol_fix_ms150_cycm <- tbl_sol_fix_ms150_cycm %>%
  filter(calvingYear > n_min_year) %>%
  filter(calvingYear <= n_max_year)
```

## Calving Year Times Month as Date
Calving year and calving month which are in separate columns are combined into a 
single column and are converted into a `Date` as data-type. This makes the use of 
a `data_scale` on the x-axis.

```{r combine-calving-year-month}
tbl_sol_fix_mar_cycm <- tbl_sol_fix_mar_cycm %>%
  mutate(`Calving Year x Month` = as.Date(paste(ifelse(nchar(calvingMonth) == 1,
                                               paste(calvingYear, paste0("0",calvingMonth), sep = '-'),
                                               paste(calvingYear, calvingMonth, sep = '-')),
                                      "01", sep = "-")))
tbl_sol_fix_hsa150_cycm <- tbl_sol_fix_hsa150_cycm %>%
  mutate(`Calving Year x Month` = as.Date(paste(ifelse(nchar(calvingMonth) == 1,
                                                       paste(calvingYear, paste0("0",calvingMonth), sep = '-'),
                                                       paste(calvingYear, calvingMonth, sep = '-')),
                                                "01", sep = "-")))
tbl_sol_fix_ms150_cycm <- tbl_sol_fix_ms150_cycm %>%
  mutate(`Calving Year x Month` = as.Date(paste(ifelse(nchar(calvingMonth) == 1,
                                                       paste(calvingYear, paste0("0",calvingMonth), sep = '-'),
                                                       paste(calvingYear, calvingMonth, sep = '-')),
                                                "01", sep = "-")))
#head(tbl_sol_fix_mar_cycm,15)
# tbl_sol_fix_mar_cycm %>% select(Solution:`Calving Year x Month`) %>% head(15)
```

## Levels in Chronological Order
The order of the levels should be sorted chronologically.

```{r }
# sort-levels
tbl_sol_fix_mar_cycm <- tbl_sol_fix_mar_cycm %>%
  arrange(desc(Level))
#head(tbl_sol_fix_mar_cycm)
tbl_sol_fix_hsa150_cycm <- tbl_sol_fix_hsa150_cycm %>%
  arrange(desc(Level))
tbl_sol_fix_ms150_cycm <- tbl_sol_fix_ms150_cycm %>%
  arrange(desc(Level))
```

The levels of the cycm effects should be unique.

```{r check-levels-unique}
tbl_sol_fix_mar_cycm %>%
  group_by(`Calving Year x Month`) %>%
  summarise(cnt = n()) %>%
  filter(cnt > 1)
tbl_sol_fix_hsa150_cycm %>%
  group_by(`Calving Year x Month`) %>%
  summarise(cnt = n()) %>%
  filter(cnt > 1)
tbl_sol_fix_ms150_cycm %>%
  group_by(`Calving Year x Month`) %>%
  summarise(cnt = n()) %>%
  filter(cnt > 1)
```

The first six records of the prepared data-set are shown below.

```{r head-data}
head(tbl_sol_fix_mar_cycm)
head(tbl_sol_fix_hsa150_cycm)
head(tbl_sol_fix_ms150_cycm)
```

# Plots
The plots show the solutions for the different levels of fixed effect. Because the 
effect-levels are time-dependent, the x-axis is shown as a time scale.

```{r }
# plots using date on x-axis
p_cycm_mar <- ggplot(data = tbl_sol_fix_mar_cycm, aes(x = `Calving Year x Month`, y=Solution)) + 
  geom_point() + 
  geom_line(color = "red") + 
  scale_x_date(name = "Calving Year X Month",
               date_breaks = "1 year", 
               date_labels = "%Y"
               ) + 
  ggtitle(label = "Merkmal: mar")
p_cycm_mar
ggsave(filename = 'solfix_cycm_mar.pdf', plot = p_cycm_mar)
ggsave(filename = 'solfix_cycm_mar.png', plot = p_cycm_mar)
ggsave(filename = 'solfix_cycm_mar.svg', plot = p_cycm_mar)

p_cycm_hsa150 <- ggplot(data = tbl_sol_fix_hsa150_cycm, aes(x = `Calving Year x Month`, y=Solution)) + 
  geom_point() + 
  geom_line(color = "red") + 
  scale_x_date(name = "Calving Year X Month",
               date_breaks = "1 year", 
               date_labels = "%Y"
  ) + 
  ggtitle(label = "Merkmal: hsa150")
p_cycm_hsa150
ggsave(filename = 'solfix_cycm_hsa150.pdf', plot = p_cycm_hsa150)
ggsave(filename = 'solfix_cycm_hsa150.png', plot = p_cycm_hsa150)
ggsave(filename = 'solfix_cycm_hsa150.svg', plot = p_cycm_hsa150)

p_cycm_ms150 <- ggplot(data = tbl_sol_fix_ms150_cycm, aes(x = `Calving Year x Month`, y=Solution)) + 
  geom_point() + 
  geom_line(color = "red") + 
  scale_x_date(name = "Calving Year X Month",
               date_breaks = "1 year", 
               date_labels = "%Y"
  ) + 
  ggtitle(label = "Merkmal: ms150")
p_cycm_ms150
ggsave(filename = 'solfix_cycm_ms150.pdf', plot = p_cycm_ms150)
ggsave(filename = 'solfix_cycm_ms150.png', plot = p_cycm_ms150)
ggsave(filename = 'solfix_cycm_ms150.svg', plot = p_cycm_ms150)
```

---
# Old Experiments and Material Taken from ggplot2 (@Wickham2009)

```{r }
# examples from the ggplot2 book
# p <- qplot(cyl, wt, data = mtcars)
# p
# p + scale_x_continuous(breaks = c(5.5, 6.5))


# use dates on x.axis
# data("economics")
# str(economics)
# plot <- qplot(date, psavert, data = economics, geom = "line") + ylab("Personal savings rate") +
#   geom_hline(yintercept = 0, colour = "grey50")
# plot
# plot + scale_x_date(
#   limits = as.Date(c("2004-01-01", "2005-01-01")),
#   date_labels = "%Y-%m-%d")
# 
# # alternative formulation
# plot2 <- ggplot(data = economics, aes(x = date, y = psavert)) + 
#   geom_line(colour = "grey50") + 
#   geom_hline(yintercept = 0)
# plot2
# 
# # limiting x-axis range when variable on x-axis is a date
# plot3 <- ggplot(data = economics, aes(x = date, y = psavert)) + 
#   geom_line(colour = "grey50") + 
#   geom_hline(yintercept = 0)
# plot3 + scale_x_date(
#   date_breaks = "1 year",
#   limits = as.Date(c("2000-01-01", "2010-01-01")), 
#   date_labels = "%Y")
```

